---
title: "CN_Analysis"
author: "Sophia"
date: "2024-04-11"
output: html_document
---


setwd("/Users/sophiafatima/Documents/7. MMM_Article/THESIS/Soil Analysis/CN Analysis") #ensure this path leads to your working directory

load("CN_Analysis.RData")

#Non-Averaged Data Analysis
```{r}
setwd("/Users/sophiafatima/Library/CloudStorage/OneDrive-MiddleburyCollege/Classes/THESIS/Soil Analysis/CN Analysis")
#setwd("C:/Users/hamia/OneDrive - Middlebury College/Classes/THESIS/Soil Analysis/CN Analysis")
library(readr)
cn <- read.csv("cn_analysis.csv")

library(dplyr)
library(ggplot2)

#Levene's test for homogeneity of variance
#This assumes we are doing a two-sample t-test
#this tests the null hypothesis that variance IS EQUAL
library(car)
leveneTest(C_conc_mgG~River,data=cn) #P=0.1121, so variance is equal

#let's do a t-test to compare means between two rivers, E & W
#w/o taking into account the different “sites”: 7 samples per site
#make sure you include the data argument
#you cannot pipe something into the t.test function :(
t.test(C_conc_mgG~River,var.equal=T,data=cn) #p-value = 0.01226

#testing variance of nitrogen
leveneTest(N_conc_mgG~River,data=cn) #P=0.1174

t.test(N_conc_mgG~River,var.equal=T,data=cn) #p-value = 0.00913

#since there was a significant difference between the rivers, I cannot do statistical test between Bay, RW, RL for both rivers at once

#Englesby
#we first construct a linear model (lm) object
cn_E <- filter(cn, River=="E")
lm.E<-lm(C_conc_mgG~Site,data=cn_E)

#next we extract the residuals of the lm object
#we will check that these are normally distributed
#residual takes each value of each sample and subtracts it from that group's mean
residuals<-residuals(lm)
shapiro.test(residuals)
#the p-value is 0.003299, so the data DO NOT follow a normal distribution

hist(residuals, data=cn_E)
#It looks like distribution is right-skewed so we take log
lm.2<-lm(log(C_conc_mgG)~Site, data=cn_E)
residuals2<-residuals(lm.2)
shapiro.test(residuals2) #p=0.1442, so the data now DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(C_conc_mgG~Site,data=cn_E)
#p=0.5779, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.2)
#p=0.01289, null hypothesis is rejected, P is significant

##Now test for nitrogen

#First I need to filter for when data is NA
cn_E_Nitrogen<-cn_E %>%
  filter(N_conc_mgG!="UNKN", N_conc_mgG>0) #removing NA & values less than 0
#removed 4 samples: EBD1, EURW2, ED1RW1, ED1RW3

lm.type<-lm(N_conc_mgG~Site,data=cn_E_Nitrogen)

residuals.type<-residuals(lm.type)
shapiro.test(residuals.type)
#the p-value is 0.1065, so the data DO follow a normal distribution, after removing 0s

hist(residuals.type, data=cn_E_Nitrogen)

#we can also test for homogeneity of variance
leveneTest(N_conc_mgG~Site,data=cn_E_Nitrogen)
#p=0.2466, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.type)

#No significant difference in Nitrogen between sites

```
#	Two Sample t-test (C)

data:  Carbon_mg by River
t = 2.6444, df = 43, p-value = 0.01138
alternative hypothesis: true difference in means between group E and group W is not equal to 0
95 percent confidence interval:
 0.09185094 0.68205103
sample estimates:
mean in group E mean in group W 
      0.6836915       0.2967405 
      
#	Two Sample t-test (N)

data:  Nitrogen_mg by River
t = 2.7519, df = 43, p-value = 0.00864
alternative hypothesis: true difference in means between group E and group W is not equal to 0
95 percent confidence interval:
 0.005493483 0.035631635
sample estimates:
mean in group E mean in group W 
     0.03442118      0.01385862 
     

#Analysis of Variance Table for log-transformed Carbon

Response: log(Carbon_mg)
          Df  Sum Sq Mean Sq F value  Pr(>F)  
Site       2  5.8186 2.90929  5.3913 0.01289 *
Residuals 21 11.3323 0.53963                  
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# Analysis of Variance Table for Nitrogen

Response: Nitrogen_mg
          Df    Sum Sq   Mean Sq F value Pr(>F)  
Site       2 0.0041238 0.0020619  3.1203 0.0701 .
Residuals 17 0.0112336 0.0006608                 
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#Tukey Test for Carbon, Englesby

```{r}
#the function for this is TukeyHSD() and it must have an lm object input into it
#you need to install and load the 'emmeans' package for this
install.packages("emmeans")
library(emmeans)
countpairs <- emmeans(lm.2, "Site")
countUnplanned <- contrast(countpairs, method = "pairwise", adjust = "tukey")
countUnplanned

#looks like sig difference between B & RL, almost significant diff between RL & RW (p=0.0682)
```
#Results of Tukey Test
 contrast estimate    SE df t.ratio p.value
 B - RL     -1.198 0.387 21  -3.094  0.0146
 B - RW     -0.378 0.387 21  -0.977  0.5990
 RL - RW     0.820 0.346 21   2.367  0.0682

Results are given on the log (not the response) scale. 
P value adjustment: tukey method for comparing a family of 3 estimates 

#Repeat everything above with Winooski Samples
```{r}
#Winooski
#we first construct a linear model (lm) object
cn_W <- filter(cn, River=="W")

##CARBON
lm.W<-lm(C_conc_mgG~Site,data=cn_W)

#next we extract the residuals of the lm object
#we will check that these are normally distributed
#residual takes each value of each sample and subtracts it from that group's mean
residuals<-residuals(lm.W)
shapiro.test(residuals)
#the p-value is 0.001653, so the data DO NOT follow a normal distribution

#check histogram
hist(residuals, data=cn_W)

#It looks like distribution is right-skewed so we take log
lm.W2<-lm(log(C_conc_mgG)~Site, data=cn_W)
residuals2<-residuals(lm.W2)
shapiro.test(residuals2) #p=0.3096, so the data now DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(C_conc_mgG~Site,data=cn_W)
#p=0.8511, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.W2)

##NITROGEN
#It seems there are too many N values below detection limit...
#If I get rid of all values of 0, I am only left w/mostly WU & WB sites
#Only 9 samples have N values not 0
#Perhaps not worth the analysis

```
#Analysis of Variance Table for log-transformed Carbon

Response: log(Carbon_mg)
          Df Sum Sq Mean Sq F value   Pr(>F)   
Site       2 16.130  8.0648  7.4964 0.004282 **
Residuals 18 19.365  1.0758                    
---
Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

#Tukey Test for Carbon, Winooski
```{r}
#the function for this is TukeyHSD() and it must have an lm object input into it
#you need to install and load the 'emmeans' package for this
library(emmeans)
countpairs <- emmeans(lm.W2, "Site")
countUnplanned <- contrast(countpairs, method = "pairwise", adjust = "tukey")
countUnplanned

#looks like sig difference between B & RL(p=0.0031) and B & RW (p=0.0197)
#no sig diff between RL & RW
```


#Visualization
```{r}
#making a box-plot
cn_E%>%
  ggplot(aes(x=Site, y=C_conc_mgG))+
  geom_boxplot()+
  xlab("Englesby Sites")+
  ylab("Carbon (mg)")+
  theme_classic()

cn_W%>%
  ggplot(aes(x=Site, y=Carbon_mg))+
  geom_boxplot()+
  xlab("Winooski Sites")+
  ylab("Carbon (mg)")+
  theme_classic()

#Making scatterplot with Winooski data
  W_C.plot<-ggplot() + geom_point(data=cn_W, aes(x = Site, y = Carbon_mg, color=name, shape=name), size=5, show.legend=TRUE)+
  scale_shape_manual(values=c(0, 1, 2, 4, 5, 6, 7))+
  labs(y = "Carbon (mg/g dry wt)", x=NULL) + 
  theme_bw() + theme(panel.border = element_blank())+
  theme(text=element_text(size=12),
        axis.text.x = element_text(angle = 45, hjust=1 ))+
    labs(color="Sampling Site", shape="Sampling Site", size=NULL)
#Making scatterplot with Englesby data
  E_C.plot<-  ggplot() + geom_point(data=cn_E, aes(x = Site, y = Carbon_mg, color=name, shape=name), size=5, show.legend=TRUE)+
  scale_shape_manual(values=c(0, 1, 2, 4, 5, 6, 7))+
  labs(y = "Carbon (mg/g dry wt)", x=NULL) + 
  labs(color="Sampling Site", shape="Sampling Site", size=NULL)+
  theme_bw() + theme(panel.border = element_blank())+
  theme(text=element_text(size=12),
        axis.text.x = element_text(angle = 45, hjust=1 ))
  
ggsave("W_C.tiff", plot = W_C.plot, dpi = 1200, width = 8, height = 5, units = "in")
ggsave("E_C.tiff", plot = E_C.plot, dpi = 1200, width = 8, height = 5, units = "in")
    
  ##NITROGEN
#Making scatterplot with Winooski data
  W_N<-ggplot() + geom_point(data=cn_W, aes(x = Site, y = Nitrogen_mg, color=name, shape=name), size=5, show.legend=TRUE)+
  scale_shape_manual(values=c(0, 1, 2, 4, 5, 6, 7))+
  labs(y = "Nitrogen (mg/g dry wt)", x=NULL) + 
  theme_bw() + theme(panel.border = element_blank())+
  theme(text=element_text(size=12),
        axis.text.x = element_text(angle = 45, hjust=1 ))+
    labs(color="Sampling Site", shape="Sampling Site", size=NULL)
#Making scatterplot with Englesby data
  E_N<-ggplot() + geom_point(data=cn_E, aes(x = Site, y = N_conc_mgG, color=name, shape=name), size=5, show.legend=TRUE)+
  scale_shape_manual(values=c(0, 1, 2, 4, 5, 6, 7))+
  labs(y = "Nitrogen (mg/g dry wt)", x=NULL) + 
  labs(color="Sampling Site", shape="Sampling Site", size=NULL)+
  theme_bw() + theme(panel.border = element_blank())+
  theme(text=element_text(size=12),
        axis.text.x = element_text(angle = 45, hjust=1 ))
  
ggsave("W_N.tiff", plot = W_C.plot, dpi = 1200, width = 8, height = 5, units = "in")
ggsave("E_N.tiff", plot = E_C.plot, dpi = 1200, width = 8, height = 5, units = "in")
```

#CONCLUSIONS:
Sig diff in carbon between sites in both Winooski & Englesby River, no sig difference in N in Englesby
Did not do statistical analysis on Englesby samples because there were too many 0 values

#REDO EVERYTHING BUT AVERAGED

```{r}
setwd("/Users/sophiafatima/Library/CloudStorage/OneDrive-MiddleburyCollege/Classes/THESIS/Soil Analysis/CN Analysis")
#setwd("C:/Users/hamia/OneDrive - Middlebury College/Classes/THESIS/Soil Analysis/CN Analysis")
library(readr)
cn <- read.csv("cn_analysis.csv")

library(dplyr)
#Taking the average of each site
###AVERAGING CN###
cn_avg<-cn%>%
    group_by(name, Site)%>%
    mutate(avg_c=mean(C_conc_mgG),
           avg_n=mean(N_conc_mgG),
          )%>%
    select(name, River, Site, avg_c, avg_n)%>%
    distinct()

library(car)
leveneTest(avg_c~River,data=cn_avg) #P=0.3654, so variance is equal

#let's do a t-test to compare means between two rivers, E & W
t.test(avg_c~River,var.equal=T,data=cn_avg) #p-value = 0.1009

#mean in group E mean in group W 
#    0.024256063     0.009970067 

#NO SIG DIFF BETWEEN RIVERS SO CAN COMBINE THE DATA

#testing variance of nitrogen
leveneTest(avg_n~River,data=cn_avg) #P=0.2223

t.test(avg_n~River,var.equal=T,data=cn_avg) #p-value = 0.09877

#mean in group E mean in group W 
#   0.0012280731    0.0004652186

#since there was not a significant difference between the rivers, I can do statistical test between Bay, RW, RL for both rivers at once

```

```{r}

### STATISTICAL ANALYSIS WITH ALL AVERAGED SAMPLES (WINOOSKI & ENGLESBY) ###

#we first construct a linear model (lm) object
lm.all<-lm(avg_c~Site,data=cn_avg)

#next we extract the residuals of the lm object
residuals<-residuals(lm.all)
shapiro.test(residuals)
#the p-value is 0.1409, so the data DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_c~Site,data=cn_avg)
#p=0.6007, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.all)
#p=0.6684, null hypothesis supported

#          Df     Sum Sq    Mean Sq F value Pr(>F)
#Site       2 1.6785e-06 8.3927e-07   0.418 0.6684
#Residuals 11 2.2086e-05 2.0078e-06  

cn_RLRW<- cn_avg%>%
  filter(Site != "B")

leveneTest(avg_n~Site,data=cn_RLRW) #P=0.4877, so variance is equal

#let's do a t-test to compare means between two rivers, E & W
t.test(avg_c~Site,var.equal=T,data=cn_RLRW) #p-value = 0.5057

#mean in group RL mean in group RW 
#      0.01940209       0.01260400 


summarized_cn<-cn_avg%>%
  group_by(River, Site)%>%
  summarize(mean(avg_n), mean(avg_c), sd(avg_n), sd(avg_c))

write.csv(summarized_cn, file="summarized_cn2.csv")

#mean in group RL mean in group RW 
#       0.5833999        0.3785860 

##Now test for nitrogen

#First I need to filter for when data is NA
cn_avg_filtered<-cn_avg %>%
  filter(avg_n>0) #removing WD1

lm.n<-lm(avg_n~Site,data=cn_avg_filtered)

residuals.n<-residuals(lm.n)
shapiro.test(residuals.n)
#the p-value is 0.07885, so the data DO NOT follow a normal distribution, after removing 0s

hist(residuals.n, data=cn_avg_filtered)
#data seem right skewed so let's take log
lm.n2<-lm(log(avg_n)~Site, data=cn_avg_filtered)
residuals2<-residuals(lm.n2)
shapiro.test(residuals2) #p=0.8537, so the data now DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_n~Site,data=cn_avg_filtered)
#p=0.7888, fail to reject null hypothesis -> variances are equal

anova(lm.n2) #p=0.2573

#No significant difference in Nitrogen between sites

### STATISTICAL ANALYSIS with SUBSETS ###
#Carbon

cn_avg.E<-cn_avg%>%
  filter(River=="E") #Englesby Brook

#we first construct a linear model (lm) object
lm.all<-lm(avg_c~Site,data=cn_avg.E)

#next we extract the residuals of the lm object
residuals<-residuals(lm.all)
shapiro.test(residuals)
#the p-value is 0.1193, so the data DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_c~Site,data=cn_avg.E)
#p=0.7862, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.all)
#p=0.4701, null hypothesis supported

#Nitrogen

#we first construct a linear model (lm) object
lm.all<-lm(avg_n~Site,data=cn_avg.E)

#next we extract the residuals of the lm object
residuals<-residuals(lm.all)
shapiro.test(residuals)
#the p-value is 0.1636, so the data DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_n~Site,data=cn_avg.E)
#p=0.7737, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.all)
#p=0.4273, null hypothesis supported


### STATISTICAL ANALYSIS with SUBSETS ###

cn_avg.W<-cn_avg%>%
  filter(River=="W")

#we first construct a linear model (lm) object
lm.all<-lm(avg_c~Site,data=cn_avg.W)

#next we extract the residuals of the lm object
residuals<-residuals(lm.all)
shapiro.test(residuals)
#the p-value is 0.2882, so the data DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_c~Site,data=cn_avg.W)
#p=0.739, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.all)
#p=0.0264, SIG DIFF

#Tukey Test
library(emmeans)
countpairs <- emmeans(lm.all, "Site")
countUnplanned <- contrast(countpairs, method = "pairwise", adjust = "tukey")
countUnplanned

 #contrast estimate      SE df t.ratio p.value
 #B - RL    0.03108 0.00707  4   4.398  0.0253
 #B - RW    0.02860 0.00707  4   4.046  0.0334
 #RL - RW  -0.00248 0.00500  4  -0.497  0.8766

#Nitrogen

#we first construct a linear model (lm) object
lm.all<-lm(avg_n~Site,data=cn_avg.W)

#next we extract the residuals of the lm object
residuals<-residuals(lm.all)
shapiro.test(residuals)
#the p-value is 0.1152, so the data DO follow a normal distribution

#we can also test for homogeneity of variance
leveneTest(avg_n~Site,data=cn_avg.W)
#p=0.7901, fail to reject null hypothesis -> variances are equal

#our assumptions are met, we can proceed to ANOVA
#the anova function requires an lm object
anova(lm.all)
#p=0.3733, null hypothesis supported

```

Analysis of Variance Table

Response: log(avg_n)
          Df Sum Sq Mean Sq F value Pr(>F)
Site       2 2.4854 1.24270  1.6116 0.2582
Residuals  8 6.1686 0.77108

#Drawing Plot with Mercury

```{r}
#write.csv(cn_avg, "/Users/sophiafatima/Documents/7. MMM_Article/THESIS/Soil Analysis/cn_avg.csv", row.names=TRUE)

setwd("/Users/sophiafatima/Documents/7. MMM_Article/THESIS/Soil Analysis/CN Analysis")

library(readr)
library(ggplot2)
cn_Hg <- read.csv("/Users/sophiafatima/Documents/7. MMM_Article/THESIS/Soil Analysis/CN Analysis/cn_hg.csv")

cn_Hg.W<-cn_Hg%>%
  filter(River=="W")

Hg_C.plot<-
cn_Hg.W%>%
  ggplot(aes(x=avg_c,y=Average_Hg))+
  geom_point()+
  theme_classic(base_size=20)+
  labs(x="Average C (mg/g dry wt)", y="Average THg (ng/g)")
  

Hg_C.plot

Hg_C.plot <- Hg_C.plot+
  geom_smooth(method="lm")

ggsave("W_Hg_C_plot.tiff", plot = Hg_C.plot, dpi = 1000, width = 8, height = 5, units = "in")

cor.test(cn_Hg.W$avg_c,cn_Hg.W$Average_Hg)

#Englesby

cn_Hg.E<-cn_Hg%>%
  filter(River=="E")

E_Hg_C.plot<-
cn_Hg.E%>%
  ggplot(aes(x=avg_c,y=Average_Hg))+
  geom_point()+
  theme_classic(base_size=20)+
  labs(x="Average C (mg/g dry wt)", y="Average THg (ng/g)")

png(file="MMM_total_class_barplot2.png", width=18, height = 10, units='in', res = 300)

E_Hg_C.plot

E_Hg_C.plot <- E_Hg_C.plot+
  geom_smooth(method="lm")

cor.test(cn_Hg.E$avg_c,cn_Hg.E$Average_Hg)
ggsave("E_Hg_C_plot.tiff", plot = E_Hg_C.plot, dpi = 1000, width = 8, height = 5, units = "in")

# Winooski N

cn_Hg.W<-cn_Hg%>%
  filter(River=="W")

Hg_N.plot<-
cn_Hg.W%>%
  ggplot(aes(x=avg_n,y=Average_Hg))+
  geom_point()+
  theme_classic(base_size=20)+
  labs(x="Average N (mg/g dry wt)", y="Average Hg (ng/g)")
  

Hg_N.plot

Hg_N.plot+
  geom_smooth(method="lm")

cor.test(cn_Hg.W$avg_n,cn_Hg.W$Average_Hg)

#Englesby N

cn_Hg.E<-cn_Hg%>%
  filter(River=="E")

Hg_N.plot<-
cn_Hg.E%>%
  ggplot(aes(x=avg_n,y=Average_Hg))+
  geom_point()+
  theme_classic(base_size=20)+
  labs(x="Average N (mg/g)", y="Average Hg (ng/g)")

Hg_N.plot

Hg_N.plot+
  geom_smooth(method="lm")

cor.test(cn_Hg.E$avg_n,cn_Hg.E$Average_Hg)
```

##ALL SAMPLES CORR TEST
	Pearson's product-moment correlation

data:  cn_Hg$avg_c and cn_Hg$Average_Hg
t = 3.5386, df = 12, p-value = 0.004081
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 0.2963885 0.9028560
sample estimates:
      cor 
0.7145928 

##WINOOSKI CORR TEST
	Pearson's product-moment correlation

data:  cn_Hg.E$avg_c and cn_Hg.E$Average_Hg
t = 3.1942, df = 5, p-value = 0.02415
alternative hypothesis: true correlation is not equal to 0
95 percent confidence interval:
 0.1727065 0.9723904
sample estimates:
      cor 
0.8192187 
